---
title: "01 - Cleaning"
output: html_notebook
---

### Reference

[Students' Adaptability Level Prediction in Online Education using Machine Learning Approaches](https://ieeexplore.ieee.org/document/9579741)

### Dataset

[Students Adaptability Level in Online Education](https://www.kaggle.com/datasets/mdmahmudulhasansuzan/students-adaptability-level-in-online-education)

#### Metadata

| Variable | Description |
| --- | --- |
| Gender | Gender type of student |
| Age | Age range of the student |
| Education Level | Education institution level |
| Institution Type | Education institution type |
| IT Student | Studying as IT student or not |
| Location | Is student location in town |
| Load-shedding | Level of load shedding |
| Financial Condition | Financial condition of family |
| Internet Type | Internet type used mostly in device |
| Network Type | Network connectivity type |
| Class Duration | Daily class duration |
| Self Lms | Institution’s own LMS availability |
| Device | Device used mostly in class |
| Adaptivity Level | Adaptability level of the student |

### Imports

```{r}
library(dplyr)
```

### Cleaning

Load dataset.

```{r}
df <- read.csv(
  "../data/students_adaptability_level_online_education.csv",
  stringsAsFactors = TRUE
)

head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```

No missing values.

Convert variables to ordered factors.

Notes:

- We will treat `Education.Level` as nominal variable to indicate the origin of the student.
- We will treat `Load.shedding` as nominal variable to indicate if the student is from high or low load shedding areas. 
- We will not convert `Adaptivity.Level` since it will be used as the dependent variable for the multiclass classification problem.

```{r}
df$Age <- df$Age %>%
  factor(
    levels = c("1-5", "6-10", "11-15", "16-20", "21-25", "26-30"),
    ordered = TRUE
  )

df$Financial.Condition <- df$Financial.Condition %>%
  factor(
    levels = c("Poor", "Mid", "Rich"),
    ordered = TRUE
  )

df$Network.Type <- df$Network.Type %>%
  factor(
    levels = c("2G", "3G", "4G"),
    ordered = TRUE
  )

df$Class.Duration <- df$Class.Duration %>%
  factor(
    levels = c("0", "1-3", "3-6"),
    ordered = TRUE
  )
```

### Descriptive Analysis

```{r}
str(df)
```

```{r}
summary(df)
```

### Exports

```{r}
saveRDS(df, "../data/clean-data.rds")
```
